curiobaby_drop model analysis

### 2021-02-01 analysis/model_analysis_v2.Rmd
final model explorations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
require(here)

@@

```{r, load-data, echo=F, FIT_MODELS=F}

load_data <- function(fname) {
  raw <- read.table(here(paste0("models/",fname,".csv")), sep=',')
  cnames <- raw[1,]
  mdat <- raw[2:nrow(raw),]
  names(mdat) = cnames
  for(c in 1:(ncol(mdat)-3)) {
    mdat[,c] = as.numeric(mdat[,c])
    if(min(mdat[,c]) < 0) mdat[,c] = mdat[,c] + min(mdat[,c])
    print(paste("shifting",c,"so minimum is 0"))
  }
  mdat$trial = rep(1:20, each=2)
  # right now, each trial is 2 rows, 
  # uniquely identified with drop_object x condition
  #mdat %>% distinct(drop_object, condition)

@@

Define helper functions

@@

feat_names = setdiff(names(mdat), c("drop","target","relation","trial"))
# for each feature, want to fit beta to maximize correlation between human and model choices
#feat_df <- get_feature_df(mdat, feat_names[1])

@@

## Fit models to all trials (MSE + r objective)

@@

```{r}
load(here("models/model_data.RData"))
```

@@

knitr::kable(fit_tab, digits = 3)
```

@@

```{r, echo=F, message=F}
require(ggpubr)

plot_model_preds <- function(feat_name, fits, hum_data, age="adults", mdat) {
  this_dat = subset(fits, Age==age & feature==feat_name)
  md <- get_model_target_choice_props(get_feature_df(mdat, feat_name), hum_data, beta)
  md %>% ggplot(aes(x=chose_target, y=Model_targ_prop, color=relation)) +
    geom_abline(slope=1, intercept=0, linetype="dashed") +
    geom_point(alpha=.5) + geom_smooth(method='lm', formula=y ~ x) + 
    xlab("Human Choice Probability") + 
    ylab("Model Choice Probability") +
    ggtitle(paste0(feat_name,": ",age)) + theme_minimal() + 
    xlim(0,1) + ylim(0,1) +
    annotate("text", x=.5, y=0.05, 
             label = paste0("beta = ",round(this_dat$beta,2),", MSE = ",round(this_dat$mse, 3))) # 
}
```

## Plots of other best-fitting features for all trials (MSE fits)

```{r, echo=F, message=F}
ad_supp <- plot_model_preds("support_probability", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch_supp <- plot_model_preds("support_probability", all_trials_mse, child_trial_agg, age="children", mdat)

ggarrange(ad_supp, ch_supp, nrow=1, common.legend = T)
```

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_combo", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_combo", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r, echo=F, message=F}
ad <- plot_model_preds("support_std", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_std", all_trials_mse, child_trial_agg, age="children", mdat)

@@

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("len_inverse_sharpe_ratio", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("len_inverse_sharpe_ratio", all_trials_mse, child_trial_agg, age="children", mdat)

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_response_linearity_r", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_linearity_r", all_trials_mse, child_trial_agg, age="children", mdat)

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_sharpness_mean", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_sharpness_mean", all_trials_mse, child_trial_agg, age="children", mdat)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
require(here)
require(kableExtra)

@@

```{r, load-data, echo=F, eval=FIT_MODELS}

load_data <- function(fname) {
  raw <- read.table(here(paste0("models/",fname,".csv")), sep=',')
  cnames <- raw[1,]
  mdat <- raw[2:nrow(raw),]
  names(mdat) = cnames
  for(c in 1:(ncol(mdat)-3)) {
    mdat[,c] = as.numeric(mdat[,c])
    if(min(mdat[,c]) < 0) mdat[,c] = mdat[,c] + min(mdat[,c])
    #print(paste("shifting",c,"so minimum is 0"))
  }
  mdat$trial = rep(1:20, each=2)
  # right now, each trial is 2 rows, 
  # uniquely identified with drop_object x condition
  #mdat %>% distinct(drop_object, condition)

@@

And define helper functions

@@

@@

## Fit models to all trials (MSE - r objective)

@@

```{r load-model-data, echo=F}
load(here("models/model_data.RData"))

feat_names = setdiff(names(mdat), c("drop","target","relation","trial"))
# for each feature, want to fit beta to maximize correlation between human and model choices
#feat_df <- get_feature_df(mdat, feat_names[1])
```

## Feature correlations

Some of these features are perfectly correlated and even have identical values: e.g. support_probability and support_combo.
All of the support_response_sharpness features with C<=.01 (C=.1 and above behave differently).
All of the support_response_sharpness_accuracy features with C<=.01 (C=.1 and above behave differently).
We remove some of these from the below tables.

```{r, echo=F, include=F}
get_feature_cor <- function(feat1, feat2, mdat) {
  supp_prob <- get_feature_df(mdat, feat1) %>%
    pivot_longer(3:12, names_to="target", values_drop_na=T)
  supp_combo <- get_feature_df(mdat, feat2) %>%
    pivot_longer(3:12, names_to="target", values_drop_na=T)
  return(list(r=cor(supp_combo$value, supp_prob$value), SSE=sum((supp_combo$value - supp_prob$value)^2)))
}

get_feature_cor("support_probability", "support_combo", mdat)
# same (1)

# support_response_sharpness_C=0.01 or C=.001 almost as good for children
get_feature_cor("support_probability", "support_response_sharpness_C=0.01", mdat)
# .91

get_feature_cor("support_response_sharpness_C=0.001", "support_response_sharpness_C=0.01", mdat) # 1 
get_feature_cor("support_response_sharpness_C=0.001", "support_response_sharpness_C=0.0001", mdat) # 1
get_feature_cor("support_response_sharpness_C=0.001", "support_response_sharpness_C=1e-05", mdat) # 1

get_feature_cor("support_response_sharpness_C=0.1", "support_response_sharpness_C=0.01", mdat) # .95
get_feature_cor("support_response_sharpness_C=0.1", "support_response_sharpness_C=1", mdat) # .94

get_feature_cor("support_response_sharpness_accuracy_C=0.001", "support_response_sharpness_accuracy_C=0.01", mdat) # 1
get_feature_cor("support_response_sharpness_accuracy_C=0.0001", "support_response_sharpness_accuracy_C=0.01", mdat)

# remove some of the identical features
all_trials_mse <- all_trials_mse %>% filter(feature!="support_combo",
                                            feature!="support_sharpness_mean",
                                            feature!="support_response_sharpness_C=0.0001", # keep .01
                                            feature!="support_response_sharpness_C=0.001",
                                            feature!="support_response_sharpness_C=1e-05",
                                            feature!="support_response_sharpness_accuracy_C=0.0001",
                                            feature!="support_response_sharpness_accuracy_C=0.001",
                                            feature!="support_response_sharpness_accuracy_C=1e-05")
all_trials_mse_r <- all_trials_mse_r %>% filter(feature!="support_combo",
                                            feature!="support_sharpness_mean",
                                            feature!="support_response_sharpness_C=0.0001", # keep .01
                                            feature!="support_response_sharpness_C=0.001",
                                            feature!="support_response_sharpness_C=1e-05",
                                            feature!="support_response_sharpness_accuracy_C=0.0001",
                                            feature!="support_response_sharpness_accuracy_C=0.001",
                                            feature!="support_response_sharpness_accuracy_C=1e-05")
supp_trials_mse <- supp_trials_mse %>% filter(feature!="support_combo",
                                            feature!="support_sharpness_mean",
                                            feature!="support_response_sharpness_C=0.0001", # keep .01
                                            feature!="support_response_sharpness_C=0.001",
                                            feature!="support_response_sharpness_C=1e-05",
                                            feature!="support_response_sharpness_accuracy_C=0.0001",
                                            feature!="support_response_sharpness_accuracy_C=0.001",
                                            feature!="support_response_sharpness_accuracy_C=1e-05")
cont_trials_mse <- cont_trials_mse %>% filter(feature!="support_combo",
                                            feature!="support_sharpness_mean",
                                            feature!="support_response_sharpness_C=0.0001", # keep .01
                                            feature!="support_response_sharpness_C=0.001",
                                            feature!="support_response_sharpness_C=1e-05",
                                            feature!="support_response_sharpness_accuracy_C=0.0001",
                                            feature!="support_response_sharpness_accuracy_C=0.001",
                                            feature!="support_response_sharpness_accuracy_C=1e-05")
```

We have `r length(feat_names)` features from 250 simulated drops per combination to optimize softmax betas for.
The tables below show fits and betas for the best-fitting 10 features (sorted by adult fit).

@@

knitr::kable(fit_tab[1:10,], digits = 3)
```

@@

```{r plot-model-vs-human, echo=F, message=F}
require(ggpubr)

plot_model_preds <- function(feat_name, fits, hum_data, age="adults", mdat) {
  this_dat = subset(fits, Age==age & feature==feat_name)
  md <- get_model_target_choice_props(get_feature_df(mdat, feat_name), hum_data, this_dat$beta)
  md %>% ggplot(aes(x=chose_target, y=Model_targ_prop, color=relation)) +
    geom_abline(slope=1, intercept=0, linetype="dashed") +
    geom_point(alpha=.5) + geom_smooth(method='lm', formula=y ~ x) + 
    xlab("Human Choice Probability") + 
    ylab("Model Choice Probability") +
    ggtitle(paste0(age,": ",feat_name)) + theme_minimal() + 
    xlim(0,1) + ylim(0,1) +
    annotate("text", x=.5, y=0.05, 
             label = paste0("beta = ",round(this_dat$beta,2),", MSE = ",round(this_dat$mse, 3))) # 
}
```

## Best feature from fitting all trials (MSE)

When fit with MSE, the best-fitting feature for both children and adults is support probability, shown below with betas optimized separately for adults' and children's data.
Note that although the MSE is quite low, the correlation between model and human responding for each trial is negative.

```{r}
ch_feat <- all_trials_mse %>% filter(Age=="children") %>%
  arrange(mse)

ad_feat <- all_trials_mse %>% filter(Age=="adults") %>%
  arrange(mse)

supp_std_prob_cor <- get_feature_cor("support_std", "support_probability", mdat) # .72

supp_resp_sharp_prob_cor <- get_feature_cor("support_response_sharpness_C=0.01", "support_probability", mdat)

ad <- plot_model_preds("support_probability", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_probability", all_trials_mse, child_trial_agg, age="children", mdat)

@@

The next-best feature for adults was support_std, which has r=`r round(supp_std_prob_cor$r, 2)` with the support_probability feature.
The next-best feature for children was support_response_sharpness_C=0.01 (or C=.001, .0001, or 1e-05), which has r=`r round(supp_resp_sharp_prob_cor$r, 2)` with the support_probability feature.
These second-best features are shown below.

```{r, echo=F, message=F}
ad <- plot_model_preds("support_std", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=0.01", all_trials_mse, child_trial_agg, age="children", mdat)

@@

# Best feature from fitting all trials (MSE - r)

When fit with a combined MSE and r objective, the best-fitting feature for both children and adults is support_response_sharpness_C=0.01, shown below with betas optimized separately for adults' and children's data.

```{r}
ch_feat <- all_trials_mse_r %>% filter(Age=="children") %>%
  arrange(mse) # support_response_sharpness_C=0.01, then support_response_sharpness_C=0.1

ad_feat <- all_trials_mse_r %>% filter(Age=="adults") %>%
  arrange(mse) # support_response_sharpness_C=0.01, then normed_velocity_std_after_first_collision_objects=target

supp_res_sharp_pt01_pt1 <- get_feature_cor("support_response_sharpness_C=0.01", "support_response_sharpness_C=0.1", mdat) # .95

supp_resp_sharp_norm_vel_targ <- get_feature_cor("support_response_sharpness_C=0.01", "normed_velocity_std_after_first_collision_objects=target", mdat)

ad <- plot_model_preds("support_response_sharpness_C=0.01", all_trials_mse_r, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=0.01", all_trials_mse_r, child_trial_agg, age="children", mdat)

The next-best feature for adults was normed_velocity_std_after_first_collision_objects=target, which has r=`r round(supp_resp_sharp_norm_vel_targ$r, 2)` with the support_response_sharpness_C=0.01 feature.
The next-best feature for children was support_response_sharpness_C=0.1, which has r=`r round(supp_res_sharp_pt01_pt1$r, 2)` with the support_response_sharpness_C=0.01 feature.
These second-best features are shown below.

```{r, echo=F, message=F}
ad <- plot_model_preds("normed_velocity_std_after_first_collision_objects=target", all_trials_mse_r, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=0.1", all_trials_mse_r, child_trial_agg, age="children", mdat)

## Best features from fitting containment trials (MSE)

The best-fitting feature from fitting only containment trials was support_probability, for both children and adults:

```{r}
ch_feat <- cont_trials_mse %>% filter(Age=="children") %>%
  arrange(mse) %>% slice(1) # support_probability then support_response_sharpness_C=10.0

ad_feat <- cont_trials_mse %>% filter(Age=="adults") %>%
  arrange(mse) %>% slice(1) # support_probability then support_std (beta=10 !)

knitr::kable(bind_rows(ch_feat, ad_feat), digits=3)
```

## Best features from fitting support trials (MSE)

The best-fitting feature for support trials for children was normed_velocity_std_after_first_collision_objects, and for adults was support_response_linearity_r (although support_probability was not far behind for adults):

```{r}
ch_feat <- supp_trials_mse %>% filter(Age=="children") %>%
  arrange(mse) %>% slice(1) #

ad_feat <- supp_trials_mse %>% filter(Age=="adults") %>%
  arrange(mse) %>% slice(1) # support_probability then support_std (beta=10 !)

bind_rows(ch_feat, ad_feat) %>% kable(digits=3)
```

### Other promising features: avg_final_radius_objects=target

avg_final_radius_objects=target shows positive correlations with children's and adults' preferences, but for many of the trials (both contain and support) the model chooses the target with small probability.

### Other promising features: support_response_sharpness_C=0.01

support_response_sharpness_C=0.01 has a fairly good MSE in particular for children, and also shows a positive correlation with children's preferences on containment trials.

@@
